#Reproducible Code - Assignment 1

####Load data

```{r}
library(lattice)
a1<-read.csv("activity.csv",header=TRUE)
str(a1)

```
####Part 1 What is mean total number of steps taken per day?    
-histogram of the total number of steps taken each day  
-agg_mean: mean total number of steps taken per day  
-agg_median:median total number of steps taken per day  

```{r}
agg_sum<-aggregate(a1$steps,list(dat=a1$date), FUN=sum, na.rm=TRUE)
agg_sum

hist(agg_sum$x,main=paste("Histogram of  total number of steps taken each day"),xlab=paste("Sum of steps each day"))

agg_mean<-aggregate(a1$steps,list(dat=a1$date), FUN=mean, na.rm=TRUE)
agg_mean

agg_median<-aggregate(a1$steps,by=list(dat=a1$date), FUN=median, na.rm=TRUE)
agg_median

```
####Part 2  What is the average daily activity pattern?  
-Time Series Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
-max_steps:Interval with Max steps 

```{r plot, fig.width=5, fig.height=5}

agg_mean_ti<-aggregate(a1$steps,list(interval=a1$interval), FUN=mean, na.rm=TRUE)
plot(agg_mean_ti$interval,agg_mean_ti$x, xlab=paste("Time interval"),ylab=paste("Average number of steps across all days"), main="Time Series Plot", type="l")

max_steps<-agg_mean_ti$interval[which.max(agg_mean_ti$x)]
max_steps

```
####Part 3  Imputing Missing Values  
-num_na:Number of missing values (NA) 

Strategy  
-Merged original data set with interval means data frame (agg_mean_ti) to replace NA steps for an interval with mean for that interval. 
-assigned the mean to the NA values for that interval in the new merged dataframe
-dropped the merged "means" column to have the same number of columns as original data set but with NAs replaced with the mean.   New data frame is a1_new  

-plot histogram of total number of steps each day (new data frame, a1_new has no NA for steps)  

new_mean: New mean after imputing missing values
new_median: New median after imputing missing values
new_sum:New median after imputing missing values

Effect on mean: No difference  
Effect on median: Yes, those that showed as NA for orginal, new had a value. Others remained same  
Effect on total number of steps: Yes, sum had higher values for the those dates that had NA before. Other dates that did not have  NA before had no effect in the new data frame  

```{r}

num_na<- sum(is.na(a1))
num_na

a12=a12<-merge(a1,agg_mean_ti,"interval")
str(a12)

a12$steps[is.na(a12$steps)]=a12$x[is.na(a12$steps)]

keep<-c("steps","date","interval")
a1_new<-a12[,keep]
str(a1_new)

new_sum<-aggregate(a1_new$steps,list(dat=a1_new$date), FUN=sum)
new_sum

hist(new_sum$x,main=paste("#2 Histogram of  total number of steps taken each day"),xlab=paste("Sum of steps each day"))

new_mean<-aggregate(a1_new$steps,list(dat=a1_new$date), FUN=mean)
new_mean

new_median<-aggregate(a1_new$steps,by=list(dat=a1_new$date), FUN=median)
new_median


```
####Part 4: Are there differences in activity patterns between weekdays and weekends?

-data frame a13 has additional column/variable wknd to show weekend or weekday  
-panel plot to show average number of steps taken by weekend/weekday


```{r, fig.width=7, fig.height=6}

a13=a1_new
a13$wknd<-ifelse(!weekdays(as.Date(a13$date)) %in% c("Saturday","Sunday"),"weekday","weekend")
str(a13)

avg_wd_we<-aggregate(a13$steps,list(interval=a13$interval,wknd=a13$wknd), FUN=mean)

xyplot(avg_wd_we$x~avg_wd_we$interval|avg_wd_we$wknd, type="l", main="Average number of steps by weekday/weekend", xlab="number of intervals", ylab="Average number of steps")

```